\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{../../../meta-inf/lib/naproche}
\input{../../lib/libraries}
\input{../../lib/foundations}
\input{../../lib/set-theory}
\begin{document}
  \begin{imports}
    \begin{forthel}
      %[prove off][check off]
      [readtex \path{libraries/source/set-theory/cardinals.ftl.tex}]
      [readtex \path{cantor-schroeder-bernstein.ftl.tex}]
      %[prove on][check on]
    \end{forthel}
  \end{imports}


  \section*{Cardinal Numbers and Maps}  

  \begin{forthel}
    \begin{proposition}\printlabel{SET_THEORY_06_5513850721927168}
      Let $x, y$ be sets and $f : x \into y$ and $a \subseteq x$.
      Then $|f[a]| = |a|$.
    \end{proposition}
    \begin{proof}
      $f \restriction a$ is a bijection between $a$ and $f[a]$.
      $f[a]$ is a set.
      Hence $|a| = |f[a]|$.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}
      Let $\kappa$ be a cardinal and $x \subseteq \kappa$.
      Then $|x| \leq \kappa$.
    \end{proposition}
    \begin{proof}
      Assume $|x| > \kappa$.
      Then $\kappa \subseteq |x|$.
      Take a bijection $f$ between $|x|$ and $x$.
      Then $f \restriction \kappa$ is an injective map from $\kappa$ to $x$.
      $\id_{x}$ is an injective map from $x$ to $\kappa$.
      Hence $x$ and $\kappa$ are equinumerous (by \nameref{cantor_schroeder_bernstein}).
      Indeed $x$ is a set.
      Thus $|x| = \kappa$.
      Contradiction.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}\printlabel{SET_THEORY_06_407116133171200}
      Let $x, y$ be sets.
      Then there exists an injective map from $x$ to $y$ iff $|x| \leq |y|$.
    \end{proposition}
    \begin{proof}
      Case there exists an injective map from $x$ to $y$.
        Consider an injective map $f$ from $x$ to $y$.
        Take a bijection $g$ from $|x|$ to $x$ and a bijection $h$ from $y$ to $|y|$.
        Then $g$ is an injective map from $|x|$ to $x$ and $h$ is an injective map from $y$ to $|y|$.
        Hence $h \circ f$ is an injective map from $x$ to $|y|$.
        Thus $(h \circ f) \circ g$ is an injective map from $|x|$ to $|y|$.
        Therefore $|x|
          = ||x||
          = |((h \circ f) \circ g)[|x|]|$
        (by \printref{SET_THEORY_06_2820082336006144}, \printref{SET_THEORY_06_5513850721927168}).
        We have $((h \circ f) \circ g)[|x|] \subseteq |y|$.
        Hence $|x| \leq |y|$.
      End.

      Case $|x| \leq |y|$.
        Take a bijection $g$ from $x$ to $|x|$ and a bijection $h$ from $|y|$ to $y$.
        We have $|x| \subseteq |y|$.
        Hence $g$ is an injective map from $x$ to $|y|$.
        Take $f = h \circ g$.
        Then $f$ is an injective map from $x$ to $y$.
        Indeed $f$ is injective.
        Indeed $h$ is an injective map from $|y|$ to $y$.
      End.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{corollary}\printlabel{SET_THEORY_06_4944303633727488}
      Let $x$ be a set and $y \subseteq x$.
      Then $|y| \leq |x|$.
    \end{corollary}
    \begin{proof}
      Define $f(v) = v$ for $v \in y$.
      Then $f$ is an injective map from $y$ to $x$.
      Hence $|y| \leq |x|$.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{corollary}\printlabel{SET_THEORY_06_7746592696172544}
      Let $x, y$ be sets such that $|y| < |x|$.
      Then $x \setminus y$ is nonempty.
    \end{corollary}
    \begin{proof}
      Assume the contrary.
      Then $x \subseteq y$.
      Hence $|x| \leq |y|$.
      Contradiction.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}\printlabel{SET_THEORY_06_192336220913664}
      Let $x, y$ be nonempty sets.
      Then there exists a surjective map from $x$ onto $y$ iff $|x| \geq |y|$.
    \end{proposition}
    \begin{proof}
      Case there exists a surjective map from $x$ onto $y$.
        Consider a surjective map $f$ from $x$ onto $y$.
        Define $g(v) =$ ``choose $u \in x$ such that $f(u) = v$ in $u$'' for $v \in y$.
        Then $g$ is an injective map from $y$ to $x$.
        Indeed we can show that $g$ is injective.
          Let $v, v' \in y$.
          Assume $g(v) = g(v')$.
          Take $u \in x$ such that $f(u) = v$ and $g(v) = u$.
          Take $u' \in x$ such that $f(u') = v'$ and $g(v') = u'$.
          Then $v
            = f(u)
            = f(g(v))
            = f(g(v'))
            = f(u')
            = v'$.
        End.
        Hence $|x| \geq |y|$.
      End.

      Case $|x| \geq |y|$.
        Then we can take an injective map $f$ from $y$ to $x$.
        Then $f^{-1}$ is a bijection between $\range(f)$ and $y$.
        Consider an element $z$ of $y$.
        Define \[ g(u) =
          \begin{cases}
            f^{-1}(u) & : u \in \range(f) \\
            z         & : u \notin \range(f)
          \end{cases} \]
        for $u \in x$.
        Then $g$ is a surjective map from $x$ onto $y$.
        Indeed we can show that every element of $y$ is a value of $g$.
          Let $v \in y$.
          Then $f(v) \in \range(f)$.
          Hence $g(f(v)) = f^{-1}(f(v)) = v$.
        End.
      End.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}\printlabel{SET_THEORY_06_5843717288099840}
      Let $x, y$ be nonempty sets.
      $|x| < |y|$ iff there exists an injective map from $x$ to $y$ and there exists no surjective map from $x$ onto $y$.
    \end{proposition}
    \begin{proof}
      There exists an injective map from $x$ to $y$ and there exists no surjective map from $x$ onto $y$ iff $|x| \leq |y|$ and $|x| \ngeq |y|$ (by \printref{SET_THEORY_06_407116133171200}, \printref{SET_THEORY_06_192336220913664}).
      $|x| \leq |y|$ and $|x| \ngeq |y|$ iff $|x| \leq |y|$ and $|x| \neq |y|$.
      $|x| \leq |y|$ and $|x| \neq |y|$ iff $|x| < |y|$.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}\printlabel{SET_THEORY_06_8113916590686208}
      Let $x, y$ be sets and $f : x \to y$ and $a \subseteq x$.
      Then $|f[a]| \leq |a|$.
    \end{proposition}
    \begin{proof}
      Case $a$ is empty. Obvious.

      Case $a$ is nonempty.
        $f \restriction a$ is a surjective map from $a$ onto $f[a]$.
        $f[a]$ is nonempty.
        Hence $|f[a]| \leq |a|$ (by \printref{SET_THEORY_06_192336220913664}).
        Indeed $a$ and $f[a]$ are sets.
      End.
    \end{proof}
  \end{forthel}
\end{document}
