\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{../../../meta-inf/lib/naproche}
\input{../../lib/libraries}
\input{../../lib/foundations}
\input{../../lib/set-theory}
\begin{document}
  \begin{imports}
    \begin{forthel}
      %[prove off][check off]
      [readtex \path{libraries/source/set-theory/axioms/choice.ftl.tex}]
      [readtex \path{libraries/source/set-theory/ordering.ftl.tex}]
      [readtex \path{libraries/source/foundations/equinumerosity.ftl.tex}]
      %[prove on][check on]
    \end{forthel}
  \end{imports}


  \section*{Cardinal Numbers}

  \begin{forthel}
    \begin{definition}[id=SET_THEORY_06_8286266038681600,printid]
      Let $x$ be a set.
      The cardinality of $x$ is the ordinal $\kappa$ such that $\kappa$ is equinumerous to $x$ and every ordinal that is equinumerous to $x$ is greater than or equal to $\kappa$.
    \end{definition}

    Let $|x|$ stand for the cardinality of $x$.
  \end{forthel}

  \begin{forthel}
    \begin{definition}[id=SET_THEORY_06_6818986081648640,printid]
      A cardinal number is an ordinal $\kappa$ such that $\kappa = |x|$ for some
      set $x$.
    \end{definition}

    Let a cardinal stand for a cardinal number.
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=SET_THEORY_06_2820082336006144,printid]
      Let $\kappa$ be a cardinal.
      Then $|\kappa| = \kappa$.
    \end{proposition}
    \begin{proof}
      $\kappa$ is an ordinal that is equinumerous to $\kappa$.
      Hence $|\kappa| \leq \kappa$.
      Consider a set $x$ such that $\kappa = |x|$.
      Then $|\kappa|$ is an ordinal that is equinumerous to $x$.
      Hence $\kappa \leq |\kappa|$.
      Thus $|\kappa| = \kappa$.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=SET_THEORY_06_6920913721229312,printid]
      Let $x, y$ be sets.
      Then $x$ and $y$ are equinumerous iff $|x| = |y|$.
    \end{proposition}
    \begin{proof}
      Case $x$ and $y$ are equinumerous.
        Take a bijection $f$ between $x$ and $y$.
        Consider a bijection $g$ between $y$ and $|y|$.
        Then $g \circ f$ is a bijection between $x$ and $|y|$ (by \printref{FOUNDATIONS_08_6435206693126144}).
        Hence $x$ and $|y|$ are equinumerous.
        Thus $|y| \geq |x|$.

        $f^{-1}$ is a bijection between $y$ and $x$.
        Consider a bijection $h$ between $x$ and $|x|$.
        Then $h \circ f^{-1}$ is a bijection between $y$ and $|x|$ (by \printref{FOUNDATIONS_08_6435206693126144}).
        Hence $y$ and $|x|$ are equinumerous.
        Thus $|x| \geq |y|$.

        Therefore $|x| = |y|$.
      End.

      Case $|x| = |y|$.
        Consider a bijection $f$ between $x$ and $|x|$ and a bijection $g$
        between $|y|$ and $y$.
        Then $g \circ f$ is a bijection between $x$ and $y$.
        Hence $x$ and $y$ are equinumerous.
      End.
    \end{proof}
  \end{forthel}
\end{document}
