\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{../../../../lib/tex/naproche}
\input{../../lib/libraries}
\input{../../lib/foundations}
\input{../../lib/set-theory}
\begin{document}
  \begin{imports}
    \begin{forthel}
      %[prove off][check off]
      [readtex \path{libraries/source/set-theory/ordering.ftl.tex}]
      %[prove on][check on]
    \end{forthel}
  \end{imports}


  \section*{Transfinite Induction}

  \begin{forthel}
    \begin{theorem}\printlabel{SET_THEORY_02_8493935460614144}
      Let $\Phi$ be a class.
      Assume that for all ordinals $\alpha$ if $\Phi$ contains all $\beta \in \alpha$ then $\Phi$ contains $\alpha$.
      Then $\Phi$ contains every ordinal.
    \end{theorem}
    \begin{proof}
      Define $B = \{ x \mid x$ is a set and if $x \in \Ord$ then $x \in \Phi \}$.

      Let us show that for all sets $x$ if $B$ contains every element of $x$ that is a set then $B$ contains $x$.
        Let $x$ be a set.
        Assume that every element of $x$ that is a set is contained in $B$.

        Case $x \notin \Ord$. Trivial.

        Case $x \in \Ord$.
          Then $\Phi$ contains all ordinals less than $x$.
          Hence $\Phi$ contains $x$.
          Thus $x \in B$.
        End.
      End.

      [prover vampire]
      Hence $B$ contains every set (by \printref{FOUNDATIONS_11_2812087589928960}).
      [prover eprover]
      Thus $\Phi$ contains every ordinal.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{theorem}\printlabel{SET_THEORY_02_7892040431960064}
      Let $\Phi$ be a class.
      
      (Initial case) Assume that $\Phi$ contains $0$.
      
      (Successor step) Assume that for all ordinals $\alpha$ if $\alpha \in \Phi$ then $\succ(\alpha) \in \Phi$.
      
      (Limit step) Assume that for all limit ordinals $\lambda$ if every $\alpha \in \lambda$ is contained in $\Phi$ then $\lambda \in \Phi$.
      
      Then $\Phi$ contains every ordinal.
    \end{theorem}
    \begin{proof}
      Let us show that for all ordinals $\alpha$ if $\Phi$ contains all ordinals
      less than $\alpha$ then $\Phi$ contains $\alpha$.
        Let $\alpha$ be an ordinal.
        Then $\alpha = 0$ or $\alpha$ is a successor ordinal or $\alpha$ is a
        limit ordinal.
        Assume that $\Phi$ contains all $\beta \in \alpha$.

        Case $\alpha = 0$. Trivial.

        Case $\alpha$ is a successor ordinal.
          Take an ordinal $\beta$ such that $\alpha = \succ(\beta)$.
          Then $\beta \in \Phi$.
          Hence $\alpha \in \Phi$ (by successor step).
        End.

        Case $\alpha$ is a limit ordinal.
          Then $\beta \in \Phi$ for all ordinals $\beta$ less than $\alpha$.
          Hence $\alpha \in \Phi$ (by limit step).
        End.
      End.

      [prover vampire]
      Thus $\Phi$ contains every ordinal (by \printref{SET_THEORY_02_8493935460614144}).
      [prover eprover]
    \end{proof}
  \end{forthel}
\end{document}
