\documentclass{article}
\usepackage{../lang/naproche}
\input{../../latex/lib/license}
\addbibresource{../../latex/lib/bibliography}

\title{The Drinker Paradox}
\author{\Naproche formalization:\\[0.5em]Marcel Schütz}
\date{2024}

\begin{document}
  \pagenumbering{gobble}
  \maketitle

  The \emph{Drinker Paradox} is a principle of classical predicate 
  logic popularised by the logician Raymond Smullyan in his 1978 book
  \textit{What Is the Name of this Book?} \cite{Smullyan1978} which can be 
  stated as:

  \begin{quotation}
    \noindent There is someone in the pub such that, if he is drinking then 
    everyone in the pub is drinking.
  \end{quotation}

  \begin{forthel}
    \begin{signature*}
      A place is a notion.
      Let $p$ denote a place.
    \end{signature*}

    \begin{signature*}
      The pub is a place.
    \end{signature*}

    \begin{signature*}
      A person is a notion.
      Let $P$ denote a person.
    \end{signature*}
    
    \begin{signature*}
      $P$ is inside $p$ is a relation.
    \end{signature*}
    
    \begin{signature*}
      $P$ is drinking is a relation.
    \end{signature*}
    
    \begin{theorem*}[Drinker Paradox]\label{drinker_paradox}
      Assume that there is a person inside the pub.
      Then there is a person $P$ inside the pub such that if $P$ is drinking then every person inside the pub is drinking.
    \end{theorem*}
    \begin{proof}[by case analysis]
      Case every person inside the pub is drinking.
        Choose a person $P$ inside the pub.
        Then $P$ is drinking and every person inside the pub is drinking.
        Hence if $P$ is drinking then every person inside the pub is drinking.
      End.
    
      Case there is a person inside the pub that is not drinking.
        Consider a person $P$ inside the pub that is not drinking.
        Then if $P$ is drinking then every person inside the pub is drinking.
      End.
    \end{proof}
  \end{forthel}

  \printbibliography
  \printlicense[CcByNcSa]{2024}{Marcel Schütz}
\end{document}
