\documentclass{naproche-library}
\libinput{preamble}
\begin{document}
\IfInputref{}{\setsectionlevel{section}}
\begin{smodule}[title=Countable and Uncountable Classes]{countable-and-uncountable-classes.ftl}
\importmodule[libraries/foundations]{finite-and-infinite-classes.ftl}

\begin{sfragment}{Countably Infinite Classes}
  \begin{definition}[forthel,id=FOUNDATIONS_14_6249029537103872]
    Let $X$ be a class.
    $X$ is \emph{countably infinite} iff $X$ is equinumerous to $\Nat$.
  \end{definition}

  \begin{proposition}[forthel,id=FOUNDATIONS_14_803449379749888]
    Let $X, Y$ be classes.
    If $X$ is countably infinite and $Y$ is equinumerous to $X$ then $Y$ is countably infinite.
  \end{proposition}
  \begin{proof}[forthel]
    Assume that $X$ is countably infinite and $Y$ is equinumerous to $X$.
    Take a bijection $f$ between $\Nat$ and $X$ and a bijection $g$ between $X$ and $Y$.
    Then $g \circ f$ is a bijection between $\Nat$ and $Y$ (by \printref{FOUNDATIONS_08_6435206693126144}).
    Indeed $X, Y$ are classes.
    Hence $Y$ is countably infinite.
  \end{proof}
\end{sfragment}

\begin{sfragment}{Countable Classes}
  \begin{definition}[forthel,id=FOUNDATIONS_14_5412969443753984]
    Let $X$ be a class.
    $X$ is \emph{countable} iff $X$ is finite or $X$ is countably infinite.
  \end{definition}

  \begin{proposition}[forthel,id=FOUNDATIONS_14_4182588499427328]
    Let $X, Y$ be classes.
    If $X$ is countable and $Y$ is equinumerous to $X$ then $Y$ is countable.
  \end{proposition}
  \begin{proof}[forthel]
    Assume that $X$ is countable and $Y$ is equinumerous to $X$.
    If $X$ is finite then $Y$ is finite.
    If $X$ is countably infinite then $Y$ is countably infinite.
    Hence $Y$ is countable.
  \end{proof}
\end{sfragment}

\begin{sfragment}{Uncountable Classes}
  \begin{definition}[forthel,id=FOUNDATIONS_14_2411928395710464]
    Let $X$ be a class.
    $X$ is \emph{uncountable} iff $X$ is not countable.
  \end{definition}

  \begin{proposition}[forthel,id=FOUNDATIONS_14_1073385358491648]
    Let $X, Y$ be classes.
    If $X$ is uncountable and $Y$ is equinumerous to $X$ then $Y$ is uncountable.
  \end{proposition}
  \begin{proof}[forthel]
    Assume that $Y$ is equinumerous to $X$.
    If $Y$ is countable then $X$ is countable.
    Hence if $X$ is uncountable then $Y$ is uncountable.
  \end{proof}
\end{sfragment}
\end{smodule}
\end{document}
