\documentclass{article}
\usepackage{latex/naproche}
\usepackage[english]{babel}

\title{Maximum modulus principle}
\author{\Naproche{} Formalization: Steffen Frerix (2018),\\
Adrian De Lon (2021), Peter Koepke (2021)}
\date{}

\newcommand{\Ball}[2]{B_{#1}(#2)}
\newcommand{\image}[2]{#1^{\to}[#2]}

\begin{document}
\pagenumbering{gobble}

\maketitle


\section{Introduction}

\noindent
We formalize that the maximum modulus principle in complex analysis is
implied by the open mapping theorem and the identity theorem.
Note that we do not give standard definitions of complex analytic notions,
but we only postulate axioms used in the proof. The axioms are satisfied over
the standard complex numbers but they also have non-standard models.
The formalization checks in under one minute on a modest laptop.

We use general set theoretic preliminaries:

\begin{forthel}
  [read examples/preliminaries.ftl.tex]
\end{forthel}

\section{Real and complex numbers}

\begin{forthel}
\begin{signature}
  A complex number is a mathematical object.
\end{signature}

Let $z, w$ denote complex numbers.

\begin{definition}
  $\mathbb{C}$ is the collection of all complex numbers.
\end{definition}

\begin{signature}
  A real number is a complex number.
\end{signature}

Let $x,y$ denote real numbers.

\begin{signature}
  $|z|$ is a real number.
\end{signature}

\begin{signature}
  $x$ is positive is an atom.
\end{signature}

Let $\varepsilon, \delta$ denote positive real numbers.

\begin{signature}
  $x < y$ is an atom.
\end{signature}
Let $x > y$ stand for $y < x$.
Let $x \leq y$ stand for $x = y$ or $x < y$.

\begin{axiom}
  If $x < y$ then not $y < x$.
\end{axiom}
\end{forthel}

\section{Open balls}

\begin{forthel}
Let $z, w$ denote complex numbers.
Let $\varepsilon, \delta$ denote positive real numbers.

\begin{signature}
  Let $\varepsilon$ be a positive real number.
  $\Ball{\varepsilon}{z}$ is a subset of $\mathbb{C}$ that contains $z$.
\end{signature}

\begin{axiom}
  $|z| < |w|$ for some element $w$ of $\Ball{\varepsilon}{z}$.
\end{axiom}

Let $M$ denote a subset of $\mathbb{C}$.

\begin{definition}
  $M$ is open iff for every element $z$ of $M$ there exists $\varepsilon$ such that
  $\Ball{\varepsilon}{z}$ is a subset of $M$.
\end{definition}

\begin{axiom}
  $\Ball{\varepsilon}{z}$ is open.
\end{axiom}

\begin{signature}
  A region is an open subset of $\mathbb{C}$.
\end{signature}

\begin{signature}
  Let $M$ be a region.
  $M$ is simply connected is an atom.
\end{signature}
\end{forthel}


\section{Holomorphic functions}

\begin{forthel}
Let $z, w$ denote complex numbers.
Let $\varepsilon, \delta$ denote positive real numbers.

\begin{signature}
  A holomorphic function is a function $f$ such that $\dom(f) \subseteq \mathbb{C}$
  and $f[\dom(f)] \subseteq \mathbb{C}$.
\end{signature}

Let $f$ denote a holomorphic function.

\begin{definition}
  A local maximal point of $f$ is an element $z$ of the domain of $f$ such that there exists
  $\varepsilon$ such that $\Ball{\varepsilon}{z}$ is a subset of the domain of $f$ and
  $|f(w)| \leq |f(z)|$ for every element $w$ of $\Ball{\varepsilon}{z}$.
\end{definition}

\begin{definition}
  Let $U$ be a subset of the domain of $f$. $f$ is constant on $U$ iff there exists
  $z$ such that $f(w) = z$ for every element $w$ of $U$.
\end{definition}

Let $f$ is constant stand for $f$ is constant on the domain of $f$.

\begin{axiom}[Open Mapping Theorem]
  Assume $f$ is a holomorphic function and $\Ball{\varepsilon}{z}$ is a subset of the domain of $f$.
  If $f$ is not constant on $\Ball{\varepsilon}{z}$
  then $f[\Ball{\varepsilon}{z}]$ is open.
\end{axiom}

\begin{axiom}[Identity theorem]
  Assume $f$ is a holomorphic function and the domain of $f$ is a region.
  Assume that $\Ball{\varepsilon}{z}$ is a subset of the domain of $f$.
  If $f$ is constant on $\Ball{\varepsilon}{z}$ then $f$ is constant.
\end{axiom}

\begin{proposition}[Maximum modulus principle]
  Assume $f$ is a holomorphic function and the domain of $f$ is a region.
  If $f$ has a local maximal point then $f$ is constant.
\end{proposition}
\begin{proof}
  Let $z$ be a local maximal point of $f$.
  Take $\varepsilon$ such that
  $\Ball{\varepsilon}{z}$ is a subset of $\dom(f)$
  and $|f(w)| \leq |f(z)|$ for every element $w$ of $\Ball{\varepsilon}{z}$.

  Let us show that $f$ is constant on $\Ball{\varepsilon}{z}$.\newline
  Proof by contradiction.
    Assume the contrary.
    Then $f[\Ball{\varepsilon}{z}]$ is open.
    We can take $\delta$ such that
    $\Ball{\delta}{f(z)}$ is a subset of $f[\Ball{\varepsilon}{z}]$.
    Therefore there exists an element $w$ of $\Ball{\varepsilon}{z}$ such that
    $|f(z)| < |f(w)|$.
    Contradiction.
  Qed.

  Hence $f$ is constant.
\end{proof}
\end{forthel}

\end{document}
