\documentclass{article}
\usepackage{naproche}
\libinput{preamble}
\begin{document}
\begin{smodule}{sets}
\importmodule[meta-inf]{macros.ftl?macros}
\importmodule[meta-inf]{vocabulary.ftl?vocabulary}
\STEXexport{
  \let\oldsubseteq\subseteq
  \let\subseteq\relax
  \let\oldsubset\subset
  \let\subset\relax
  \let\oldsupseteq\supseteq
  \let\supseteq\relax
  \let\oldsupset\supset
  \let\supset\relax
  \let\oldsubsetneq\subsetneq
  \let\subsetneq\relax
  \let\oldsupsetneq\supsetneq
  \let\supsetneq\relax
  \let\oldemptyset\emptyset
  \let\emptyset\relax
  \let\oldcup\cup
  \let\cup\relax
  \let\oldcap\cap
  \let\cap\relax
  \let\oldsetminus\setminus
  \let\setminus\relax
  \let\oldtimes\times
  \let\times\relax
  \let\oldbigcup\bigcup
  \let\bigcup\relax
  \let\oldbigcap\bigcap
  \let\bigcap\relax
}
\symdef{subseteq}[op=\oldsubseteq]{\;\oldsubseteq\;}
\symdef{subset}[op=\oldsubset]{\;\oldsubset\;}
\symdef{supseteq}[op=\oldsupseteq]{\;\oldsupseteq\;}
\symdef{supset}[op=\oldsupset]{\;\oldsupset\;}
\symdef{subsetneq}[op=\oldsubsetneq]{\;\oldsubsetneq\;}
\symdef{supsetneq}[op=\oldsupsetneq]{\;\oldsupsetneq\;}
\symdef{emptyset}[op=\oldemptyset]{\oldemptyset}
\symdef{set}[args=a,op=\{\cdot\}]{\comp{\{}\argsep{#1}{\comp,}\comp{\}}}
\symdef{cup}[op=\oldcup]{\,\oldcup\,}
\symdef{cap}[op=\oldcap]{\,\oldcap\,}
\symdef{setminus}[op=\oldsetminus]{\,\oldsetminus\,}
\symdef{pow}[args=1,op=\mathcal P]{\mathcal P\dobrackets{#1}}
\symdef{times}[op=\oldtimes]{\,\oldtimes\,}
\symdef{bigcup}[op=\oldbigcup]{\oldbigcup\,}
\symdef{bigcap}[op=\oldbigcap]{\oldbigcap\,}

\section{Classes and Sets}

\begin{forthel}
  [readtex \path{meta-inf/source/vocabulary.ftl.tex}]

  [readtex \path{meta-inf/source/macros.ftl.tex}]
\end{forthel}

\begin{signature*}
  A \emph{class} is a notion.
\end{signature*}

\begin{forthel}
  Let $A,B,C$ denote classes.
\end{forthel}

\begin{signature*}
  An \emph{object} is a notion.
\end{signature*}

\begin{forthel}
  Let $a,a',b,b',c,c'$ denote objects.
\end{forthel}

\begin{definition*}
  A \emph{set} is a class that is an object.
\end{definition*}


\subsection{Subclasses and Subsets}

\begin{definition*}[forthel,id=SubclassDef,printid]
  A \emph{subclass of $A$} is a class $B$ such that every element of $B$ is an element of $A$.
  Let $B\subseteq A$ stand for $B$ is a subclass of $A$.
  Let $B\subset A$ stand for $B\subseteq A$.

  Let a \emph{superclass of $B$} stand for a class $A$ such that $B\subseteq A$.
  Let $B\supseteq A$ stand for $B$ is a superclass of $A$.
  Let $B\supset A$ stand for $B\subseteq A$.

  Let a \emph{proper subclass of $A$} stand for a subclass $B$ of $A$ such that $B\neq A$.
  Let $B\subsetneq A$ stand for $B$ is a proper subclass of $A$.

  Let a \emph{proper superclass of $B$} stand for a superclass $A$ of $B$ such that $A \neq B$.
  Let $B\supsetneq A$ stand for $B$ is a proper superclass of $A$.

  Let \emph{$A$ includes $B$} stand for $B\subseteq A$.
  Let $B$ is \emph{included in $A$} stand for $B\subseteq A$.
\end{definition*}

\begin{axiom*}[forthel,title=Class Extensionality Axiom,id=ClassExtensionalityAx,printid]
  If $A\subseteq B$ and $B\subseteq A$ then $A\eq B$.
\end{axiom*}

\begin{definition*}[forthel,id=SubsetDef,printid]
  A \emph{subset of $A$} is a subclass of $A$ that is a set.

  Let a \emph{superset of $B$} stand for a class $A$ such that $B\subseteq A$.

  Let a \emph{proper subset of $A$} stand for a subset $B$ of $A$ such that $B\neq A$.

  Let a \emph{proper superset of $B$} stand for a superset $A$ of $B$ such that $A \neq B$.
\end{definition*}

\begin{axiom*}[forthel,title=Separation Axiom,id=SeparationAx,printid]
  If $B$ is a set and $A\subseteq B$ then $A$ is a set.
\end{axiom*}


\subsection{The Empty Set}

\begin{definition*}[forthel,id=EmptyDef,printid]
  $A$ is \emph{empty} iff $A$ has no elements.
  Let $A$ is \emph{nonempty} stand for $A$ is not empty.
\end{definition*}

\begin{definition*}[forthel,id=EmptySetDef,printid]
  $\emptyset=\{x\mid x\neq x\}$.
\end{definition*}

\begin{axiom*}[forthel,title=Empty Set Axiom,id=EmptySetAx,printid]
  $\emptyset$ is a set.
\end{axiom*}


\subsection{Unordered Pairs and Singletons}

\begin{definition*}[forthel,id=UnorderedPairDef,printid]
  $\set{a,b}=\{x\mid x\eq a\text{ or }x\eq b\}$.
\end{definition*}

\begin{axiom*}[forthel,title=Pairing Axiom,id=UnorderedPairAx,printid]
  $\set{a,b}$ is a set.
\end{axiom*}

\begin{definition*}[forthel,id=SingletonDef,printid]
  $\set{a}=\{x\mid x\eq a\}$.
\end{definition*}

\begin{proposition*}[forthel,id=SingletonProp,printid]
  $\set{a}$ is a set.
\end{proposition*}


\subsection{Systems of Sets}

\begin{definition*}[forthel,id=SystemOfSetsDef,printid]
  A \emph{system of sets} is a class $X$ such that every element of $X$ is a set.
\end{definition*}

\begin{forthel}
  Let $X,Y,Z$ denote systems of sets.
\end{forthel}

\begin{definition*}[forthel,id=BigUnionDef,printid]
  $\bigcup X=\{a\mid a\in x\text{ for some  }x\in X\}$.
  Let the \emph{union over $X$} stand for $\bigcup X$.
\end{definition*}

\begin{axiom*}[forthel,title=Union Axiom,id=BigUnionAx,printid]
  If $X$ is a set then $\bigcup X$ is a set.
\end{axiom*}

\begin{definition*}[forthel,id=BigIntersectionDef,printid]
  $\bigcap X=\{a\mid a\in x\text{ for all }x\in X\}$.
  Let the \emph{intersection over $X$} stand for $\bigcap X$.
\end{definition*}

\begin{proposition*}[forthel,id=BigIntersectionProp,printid]
  If $X$ is nonempty then $\bigcap X$ is a set.
\end{proposition*}


\subsection{Union, Intersection, Complement}

\begin{definition*}[forthel,id=UnionDef,printid]
  $A\cup B=\{x\mid x\in A\text{ or }x\in B\}$.
  Let the \emph{union of $A$ and $B$} stand for $A\cup B$.
\end{definition*}

\begin{forthel}
  \begin{proposition*}[id=UnionProp,printid]
    If $A$ and $B$ are sets then $A\cup B$ is a set.
  \end{proposition*}
  \begin{proof}
    Assume that $A$ and $B$ are sets.
    Then $A\cup B\eq \bigcup\set{A,B}$.
  \end{proof}
\end{forthel}

\begin{definition*}[forthel,id=IntersectionDef,printid]
  $A\cap B=\{x\mid x\in A\text{ and }x\in B\}$.
  Let the \emph{intersection of $A$ and $B$} stand for $A\cap B$.
\end{definition*}

\begin{proposition*}[forthel,id=IntersectionProp,printid]
  If $A$ is a set or $B$ is a set then $A\cap B$ is a set.
\end{proposition*}

\begin{definition*}[forthel,id=ComplementDef,printid]
  $A\setminus B=\{x\mid x\in A\text{ and }x\notin B\}$.
  Let the \emph{complement of $B$ in $A$} stand for $A\setminus B$.
\end{definition*}

\begin{proposition*}[forthel,id=ComplementProp,printid]
  If $A$ is a set then $A\setminus B$ is a set.
\end{proposition*}

\begin{definition*}[forthel,id=DisjointDef,printid]
  $A$ and $B$ are \emph{disjoint} iff $A$ and $B$ have no common elements.
\end{definition*}


\subsection{Powerclasses and Powersets}

\begin{definition*}[forthel,id=PowerclassDef,printid]
  $\pow{A}=\{x\mid x\subseteq A\}$.
  Let the \emph{powerclass of $A$} stand for $\pow{A}$.
\end{definition*}

\begin{axiom*}[forthel,title=Powerset Axiom,id=PowersetAx,printid]
  If $A$ is a set then $\pow{A}$ is a set.
  Let the \emph{powerset of $A$} stand for $\pow{A}$.
\end{axiom*}


\subsection{Ordered Pairs and Cartesian Products}

\begin{axiom*}[forthel,title=Pair Extensionality Axiom,id=PairExtensionalityAx,printid]
  If $\pair{a}{b}\eq \pair{a'}{b'}$ then $a\eq a'$ and $b\eq b'$.
\end{axiom*}

\begin{definition*}[forthel,id=CartesianProductDef,printid]
  $A\times B=\{\pair{a}{b}\mid a\in A\text{ and }b\in B\}$.
  Let the \emph{Cartesian product of $A$ and $B$} stand for $A\times B$.
\end{definition*}

\begin{proposition*}[forthel,id=CartesianProductProp,printid]
  $\pair{a}{b}\in A\times B$ iff $a\in A$ and $b\in B$.
\end{proposition*}
\end{smodule}
\end{document}
