\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{../lib/tex/forthel}

\title{Tarski's axioms for Euclidian geometry}
\author{}
\date{}

\begin{document}
  \pagenumbering{gobble}

  \maketitle

  \begin{forthel}
    [synonym point/-s]


    % This is a formalization of the beginnings of Tarskian geometry,
    % mainly following the outline of 'Metamathematische Methoden in
    % der Geometrie' by Schwabhäuser, Szmielew and Tarski. We refer to
    % this book as 'SST' in later comments.


    \begin{signature}
      A point is a notion.
    \end{signature}

    Let x,y,z,u,v,w,p,q,r,g,h denote points.


    \begin{signature}
      Cong(x,y,v,w) is an atom.
    \end{signature}

    Let x and y are congruent to v and w stand for Cong(x,y,v,w).
    Let x-y : v-w stand for x and y are congruent to v and w.

    \begin{signature}
      Betw(x,u,y) is an atom.
    \end{signature}

    Let u is between x and y stand for Betw(x,u,y).
    Let x-u-y stand for u is between x and y.
    Let !x-u-y stand for u is not between x and y.


    \begin{definition}
      Col(x,y,z) iff x-y-z or y-z-x or z-x-y.
    \end{definition}

    Let p is colinear with x and y stand for Col(p,x,y).


    % Reflexivity of congruence.
    \begin{axiom}[A1]
      x-y : y-x.
    \end{axiom}

    % Transitivity of congruence.
    \begin{axiom}[A2]
      If x-y : z-u and x-y : v-w then z-u : v-w.
    \end{axiom}

    % Identity of congruence.
    \begin{axiom}[A3]
      If x-y : z-z then x = y.
    \end{axiom}

    % Segment construction.
    \begin{axiom}[A4]
      There exists a point z such that x-y-z and y-z : p-q.
    \end{axiom}

    % Five segments.
    \begin{axiom}[A5]
      If x != y and x-y-z and p-q-r and x-y : p-q and y-z : q-r and x-u : p-v
      and y-u : q-v then z-u : r-v.
    \end{axiom}

    % Identity of betweenness.
    \begin{axiom}[A6]
      If y is between x and x then x = y.
    \end{axiom}

    % Inner Pasch.
    \begin{axiom}[A7]
      If x-u-z and y-v-z then there exists a point w such that u-w-y and v-w-x.
    \end{axiom}

    % Lower dimension.
    \begin{axiom}[A8]
      There exist points a,b,c such that !a-b-c and !b-c-a and !c-a-b.
    \end{axiom}

    % Upper dimension.
    \begin{axiom}[A9]
      If x-u : x-v and y-u : y-v and  z-u : z-v and u != v then x-y-z and y-z-x
      and z-x-y.
    \end{axiom}

    % Euclid.
    \begin{axiom}[A10]
      If x-r-v and y-r-z and x != r then there exist points s,t such that x-y-s
      and x-z-t and s-v-t.
    \end{axiom}

    % Circle continuity axiom.
    %
    % This axiom is equivalent to the the following statement: A line that has a
    % point within a circle intersects that circle (i.e. shares a point with
    % that circle.).
    %
    \begin{axiom}[CA]
      Assume z-p-q and z-p-r and z-x : z-p and z-y : z-r. Then there exists w
      such that z-w : z-q and x-w-y.
    \end{axiom}


    % This is definition 2.10 in SST. We say that the points x,y,z,r,u,v,w,p are
    % in an outer five-segment-configuration whenever OFS(x,y,z,r,u,v,w,p).
    % Using this predicate we can restate axiom A5 in the following manner:
    %
    % A5'. If OFS(x,y,z,r,u,v,w,p) and x != y then z-r : w-p.
    %
    \begin{definition}[OFS]
      OFS(x,y,z,r,u,v,w,p) iff x-y-z and u-v-w and x-y : u-v and y-z : v-w and
      x-r : u-p and y-r : v-p.
    \end{definition}

    % Reflexivity of congruence.
    \begin{lemma}[L2_1]
      x-y : x-y.
    \end{lemma}

    % Symmetry of congruence.
    \begin{lemma}[L2_2]
      Assume x-y : v-w. Then v-w : x-y.
    \end{lemma}

    % Transitivity of congruence.
    \begin{lemma}[L2_3]
      Assume x-y : v-w and v-w : p-q. Then x-y : p-q.
    \end{lemma}

    % Cong is independent of the order of the pairs.
    \begin{lemma}[L2_4]
      Assume x-y : v-w. Then y-x : v-w.
    \end{lemma}

    \begin{lemma}[L2_5]
      Assume x-y : v-w. Then x-y : w-v.
    \end{lemma}

    % Zero-length segments are congruent.
    \begin{lemma}[L2_8]
      x-x : y-y.
    \end{lemma}

    % Concatenation of segments.
    \begin{lemma}[L2_11]
      Assume x-y-z and r-v-w and x-y : r-v and y-z : v-w. Then x-z : r-w.
    \end{lemma}
    %
    % The following proof follows the proof from SST and works, but can be
    % omitted entirely.
    %
    \begin{proof}
      OFS(x,y,z,x,r,v,w,r).     % By previous results and assumption.
      Assume x = y. Then r = v. % Axiom A3 gives this implication.
      Assume x != y.            % Axiom A5 completes the proof.
    \end{proof}

    % Uniqueness for Axiom A4.
    \begin{lemma}[L2_12]
      Assume q != x. Suppose q-x-y and x-y : v-w and q-x-r and x-r : v-w. Then
      y = r.
    \end{lemma}

    % Right-betweenness.
    \begin{lemma}[L3_1]
      x-y-y.
    \end{lemma}

    % Symmetry of betweenness.
    \begin{lemma}[L3_2]
      Assume x-y-z. Then z-y-x.
    \end{lemma}

    % Left-betweenness.
    \begin{lemma}[L3_3]
      x-x-y.
    \end{lemma}

    \begin{lemma}[L3_4]
      Assume x-y-z and y-x-z. Then x = y.
    \end{lemma}

    \begin{lemma}[L3_5a]
      Assume x-y-v and y-z-v. Then x-y-z.
    \end{lemma}

    \begin{lemma}[L3_5b]
      Assume x-y-v and y-z-v. Then x-y-v.
    \end{lemma}

    \begin{lemma}[L3_6a]
      Assume x-y-z and x-z-r. Then y-z-r.
    \end{lemma}

    \begin{lemma}[L3_6b]
      Assume x-y-z and x-z-r. Then x-y-r.
    \end{lemma}

    \begin{lemma}[L3_7a]
      Assume x-y-z and y-z-r and y != z. Then x-z-r.
    \end{lemma}
    \begin{proof}
    	Take v such that x-z-v and z-v : z-r.
    	Then y-z-v and z-v : z-r. Hence v = r.
    \end{proof}

    \begin{lemma}[L3_7b]
      Assume x-y-z and y-z-r and y != z. Then x-y-r.
    \end{lemma}

    % Existence of at least two points follows from A8.
    % (All other axioms also hold in a one-point space.)
    \begin{lemma}[L3_13]
      x != y for some x, y.
    \end{lemma}

    \begin{lemma}[L3_14]
      There exist z such that x-y-z and y != z.
    \end{lemma}


    \begin{lemma}[L3_17]
      Assume x-y-z and u-v-z and x-p-u. Then there exist q such that (p-q-z and
      y-q-v).
    \end{lemma}
    \begin{proof}
      x-p-u and z-v-u.
    	Take r such that v-r-x and p-r-z. % A7 (Pasch).
    	Take q such that r-q-z and v-q-y. % A7 (Pasch).
    \end{proof}


    % This is definition 4.1 in SST. We say that the points x,y,z,r,u,v,w,p are
    % in an inner five-segment-configuration whenever IFS(x,y,z,r,u,v,w,p).
    %
    \begin{definition}[IFS]
      IFS(x,y,z,r,v,w,p,q) iff (x-y-z and v-w-p and x-z : v-p and y-z : w-p and
      x-r : v-q and z-r : p-q).
    \end{definition}

    % We can swap x,y and v,w.
    \begin{axiom}[L4_2]
      Assume IFS(x,y,z,r,v,w,p,q). Then y-r : w-q.
    \end{axiom}

    % If we have two three-point segments, with the same total length and each
    % with a segment of the same length, then the remaining segments must also
    % have the same length.
    %
    % TODO: Write a proof for this.
    %
    \begin{axiom}[L4_3]
      Assume x-y-z and r-v-w and x-z : r-w and y-z : v-w. Then x-y : r-v.
    \end{axiom}

    \begin{definition}[L4_4]
      x-y-z : u-v-w iff x-y : u-v and x-z : u-w and y-z : v-w.
    \end{definition}

    \begin{lemma}[L4_5]
      Assume x-y-z and x-z : r-w. Then there exists v such that (r-v-w and
      x-y-z : r-v-w).
    \end{lemma}
    \begin{proof}
    	Take u such that w-r-u and r != u. Then Take v such that u-r-v and
      r-v : x-y. Take g such that u-v-g and v-g : y-z. Then x-z : r-w. Therefore
      g = w.
    \end{proof}

    \begin{lemma}[L4_6]
      Assume x-y-z and x-y-z : r-v-w. Then r-v-w.
    \end{lemma}
    \begin{proof}
    	Take u such that r-u-w and x-y-z : r-u-w.
    	Then r-u-w : r-v-w and IFS(r,u,w,u,r,u,w,v).
    	Then u-u : u-v. Hence u = v. Hence r-v-w.
    \end{proof}

    \begin{lemma}[L4_11a]
      Assume Col(x,y,z). Then Col(y,z,x).
    \end{lemma}

    \begin{lemma}[L4_11b]
      Assume Col(x,y,z). Then Col(z,x,y).
    \end{lemma}

    \begin{lemma}[L4_11c]
      Assume Col(x,y,z). Then Col(z,y,x).
    \end{lemma}

    \begin{lemma}[L4_11d]
      Assume Col(x,y,z). Then Col(y,x,z).
    \end{lemma}

    \begin{lemma}[L4_11e]
      Assume Col(x,y,z). Then Col(x,z,y).
    \end{lemma}

    \begin{lemma}[L4_12]
      Col(x,x,y).
    \end{lemma}

    \begin{lemma}[L4_13]
      Assume Col(x,y,z) and x-z : r-w and r-v-w. Then Col(r,v,w).
    \end{lemma}

    \begin{lemma}[L4_14_1]
      x-y-z : u-v-w iff y-x-z : v-u-w.
    \end{lemma}

    \begin{lemma}[L4_14_2]
      x-y-z : u-v-w iff z-y-x : w-v-u.
    \end{lemma}

    \begin{lemma}[L4_14_3]
      x-y-z : u-v-w iff x-z-y : u-w-v.
    \end{lemma}

    \begin{axiom}[L4_14]
      Assume Col(x,y,z) and x-y : r-v. Then there exists w such that
      x-y-z : r-v-w.
    \end{axiom}

    \begin{definition}[L4_15]
      FS(x,y,z,r,v,w,p,q) iff Col(x,y,z) and x-y-z : v-w-p and x-r : v-q and
      y-r : w-q.
    \end{definition}

    \begin{axiom}[L4_16]
      Assume FS(x,y,z,r,v,w,p,q) and x != y. Then z-r : p-q.
    \end{axiom}


    \begin{lemma}[L4_17]
      Assume x != y and Col(x,y,z) and x-p : x-q and y-p :  y-q. Then z-p : z-q.
    \end{lemma}
    \begin{proof}
    	FS(x,y,z,p,x,y,z,q).
    \end{proof}


    \begin{lemma}[L4_18]
      Assume x != y and Col(x,y,z) and x-z : x-p and y-z : y-p. Then z = p.
    \end{lemma}

    \begin{lemma}[L4_19]
      Assume x-z-y and x-z : x-p and y-z : y-p. Then z = p.
    \end{lemma}
    \begin{proof}
      Assume x = y. Then x = z and x = p. Hence z = p.
      Assume x != y.
    \end{proof}

    % The 11th axiom of Tarski's axiomatic system says that if x-y-w and x-z-w
    % then either x-y-z or x-z-y. To show that it follows from the first ten
    % axioms we first prove Lemma C5_1 from which we can easy deduce the 11th
    % axiom.
    %
    % The definitions, lemmata and axioms C5_1a - C5_1p are not part of SST. We
    % have opted for adding them to improve proof-checking speed and readability
    % of the text.

    \begin{definition}[C5_1a]
      Betw5(x,y,z,r,p) iff x-y-z and x-y-r and x-y-p and x-z-r and x-z-p and
      x-r-p and y-z-r and y-z-p and y-r-p and z-r-p.
    \end{definition}

    Let x~y~z~r~p stand for Betw5(x,y,z,r,p).

    % The following 4 predicates state the already proven statements for
    % different positions in the proof. They are not defined in the book
    % "Metamathematische Methoden in der Geometrie. We use them because they
    % seem to increase the performance of the proof assistant, when checked just
    % before the next proving step.

    \begin{definition}[C5_1b]
      Th(x,y,z,r,p,q,g,h) iff x != y and x-y-z and x-y-r and x-r-p and r-p : z-r
      and x-z-q and z-q : z-r and z-q-h and r-p-g.
    \end{definition}

    \begin{definition}[C5_1c]
      Th2(x,y,z,r,p,q,g) iff x != y and x~y~z~q~g and x~y~r~p~g and r-p : z-r
      and z-q : z-r and y-p : g-z and y-g : g-y.
    \end{definition}

    \begin{definition}[C5_1d]
      Th3(x,y,z,r,p,q,g,u) iff Th2(x,y,z,r,p,q,g) and OFS(y,z,q,p,g,p,r,z) and
      p-q : z-r and z-u-p and r-u-q and IFS(r,u,q,z,r,u,q,p) and
      IFS(z,u,p,r,z,u,p,q) and u-r : u-q.
    \end{definition}

    \begin{definition}[C5_1e]
      Th4(x,y,z,r,p,q,g,u,v,w,h) iff Th3(x,y,z,r,p,q,g,u) and z != p and z != q
      and p-z-v and z-v : z-q and q-z-h and z-h : z-u and v-h-w and h-w : h-v.
    \end{definition}

    % For the following 5 Statements we did not find a proof yet that gets
    % checked positive by Naproche SAD. They are all used in the proof of Lemma
    % C5_1p and Lemma C5_1.

    \begin{lemma}[C5_1f]
      Assume x != y and x-y-z and x-y-r. Then there exist points a,b such that
      x-r-a and r-a : z-r and x-z-b and z-b : z-r.
    \end{lemma}
    \begin{proof}
    	Take point a such that x-r-a and r-a : z-r (by A4).
    	Take point b such that x-z-b and z-b : z-r (by A4).
    \end{proof}

    \begin{axiom}[C5_1g]
      If Th(x,y,z,r,p,q,g,h) then x~y~z~q~h and x~y~r~p~g.
    \end{axiom}

    \begin{axiom}[C5_1h]
      Assume Th(x,y,z,r,p,q,g,h) and x~y~z~q~h and x~y~r~p~g. Then y-p : h-z.
    \end{axiom}

    \begin{axiom}[C5_1i]
      Assume Th2(x,y,z,r,p,q,g) and OFS(y,z,q,p,g,p,r,z). Then p-q : z-r.
    \end{axiom}

    \begin{axiom}[C5_1j]
      Assume Th(x,y,z,r,p,q,g,h) and x~y~z~q~h and x~y~r~p~g and y-p : h-z. Then
      y-g : h-y.
    \end{axiom}



    \begin{lemma}[C5_1k]
      Assume x != y and x-y-z and x-y-r and x-r-p and r-p : z-r and x-z-q and
      z-q : z-r and (z = p or r = q). Then x-z-r or x-r-z.
    \end{lemma}

    \begin{lemma}[C5_1l]
      Assume x != y and x-y-z and x-y-r and x-r-p and r-p : z-r and x-z-q and
      z-q : z-r. Then there exist points s,t such that z-q-t and r-p-s.
    \end{lemma}

    \begin{lemma}[C5_1m]
      Assume Th2(x,y,z,r,p,q,g). Then OFS(y,z,q,p,g,p,r,z).
    \end{lemma}

    \begin{lemma}[C5_1n]
      Assume Th2(x,y,z,r,p,q,g) and OFS(y,z,q,p,g,p,r,z) and p-q : z-r. Then
      there exist u such that z-u-p and r-u-q.
    \end{lemma}

    \begin{lemma}[C5_1o]
      Assume Th4(x,y,z,r,p,q,g,u,v,w,h). Then OFS(q,z,h,v,v,z,u,q) and h-v : u-q
      and h-w : u-r and OFS(q,u,r,z,v,h,w,z) and q-r : v-w and z-w : z-r and
      z-v : z-w.
    \end{lemma}
    \begin{proof}
    	OFS(q,z,h,v,v,z,u,q). Hence h-v : u-q. Hence h-w : u-r. % TODO: this step is slow.
    	Therefore OFS(q,u,r,z,v,h,w,z). Hence q-r : v-w.
    	If q != u then z-w : z-r. If q = u then q = r. % TODO: this step is slow.
    	Then v = w. Therefore z-w : z-r.
    	Hence z-v : z-w.
    \end{proof}

    % The Idea to proof Lemma C5_1 is to extend the line x-z and x-r through two
    % points p,q such that r-p : z-r and z-q : z-r. Then one can easy see that
    % if either z = p or r = q, x-y-z or x-z-y must hold. The following Lemma
    % proofs that if there exist such points p and q then z = p or r = q must
    % hold. To see that such points exist one has to use Axiom A4 twice.

    \begin{lemma}[C5_1p]
      Assume x != y and x-y-z and x-y-r and x-r-p and r-p : z-r and x-z-q and
      z-q : z-r. Then z = p or r = q.
    \end{lemma}
    \begin{proof}
    	Take points s,t such that z-q-t and r-p-s.
    	Then Th(x,y,z,r,p,q,s,t).
    	Then x~y~z~q~t and x~y~r~p~s.
    	Then y-p : t-z.
    	Then y-t : t-y.
    	Then s = t.
    	Then Th2(x,y,z,r,p,q,s).
    	Then OFS(y,z,q,p,s,p,r,z).
    	Then p-q : z-r.
    	Take u such that z-u-p and r-u-q (by C5_1n).
    	Then IFS(r,u,q,z,r,u,q,p).
    	Then IFS(z,u,p,r,z,u,p,q).
    	Then u-z : u-p.
    	Then u-r : u-q.
    	Assume z != p. Then z != q.
    		Take v such that p-z-v and z-v : z-q.
    		Take h such that q-z-h and z-h : z-u.
    		Take w such that v-h-w and h-w : h-v.
    		Then OFS(q,z,h,v,v,z,u,q) and h-v : u-q and h-w : u-r and
    			OFS(q,u,r,z,v,h,w,z) and q-r : v-w and z-w : z-r and z-v : z-w (by C5_1o).
    		Then h-v : h-w. z != p. Hence h != z. Col(h,z,q). Therefore q-v : q-w.
    		Then z-v : z-w. z != q. Col(z,q,y). Therefore y-v : y-w.
    		Then z-v : z-w. z != q. Col(z,q,s). Therefore s-v : s-w.
    		Then y != s.
    		Then q = r.
    	Assume z = p.
    \end{proof}

    \begin{lemma}[D5_1]
      Assume x != y and x-y-z and x-y-r. Then x-z-r or x-r-z.
    \end{lemma}
    \begin{proof}
    	Take p,q such that x-r-p and r-p : z-r and x-z-q and z-q : z-r.
    	Then z = p or r = q (by C5_1p). Therefore x-z-r or x-r-z (by C5_1k).
    \end{proof}

    \begin{lemma}[D5_2]
      Assume x != y and x-y-z and x-y-r. Then y-z-r or y-r-z.
    \end{lemma}

    \begin{theorem}[D5_3]
      If x-y-w and x-z-w then x-y-z or x-z-y.
    \end{theorem}
  \end{forthel}

\end{document}
