\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb}
\usepackage{../lib/tex/naproche}

\title{group.lean in \Naproche}
\author{Formalization: Peter Koepke}
\date{2021}

\begin{document}
  \pagenumbering{gobble}

  \maketitle

% This file is a ForTheL version of 
% https://github.com/leanprover/lean/blob/master/library/init/algebra/group.lean

\begin{forthel}

\begin{lemma} Every set is a class.
\end{lemma}

[synonym type/-s]
\begin{signature} A type is a set.
\end{signature}
Let $\alpha$ stand for a type.
Let $a : t$ stand for $a$ is an element of $t$.

\begin{signature} A type with multiplication is a type.
\end{signature}
\begin{signature} Let $\alpha$ be a type with multiplication and $a,b : \alpha$. 
$a *^{\alpha} b$ is an element of $\alpha$.
\end{signature}
  
[synonym semigroup/-s]
\begin{definition} A semigroup is a type with multiplication $\alpha$
such that for all $a,b,c : \alpha$
$(a *^{\alpha} b) *^{\alpha} c = a *^{\alpha} (b *^{\alpha} c)$. 
\end{definition}

\begin{definition} A commutative semigroup is a semigroup $\alpha$
such that for all $a,b : \alpha$
$a *^{\alpha} b = b *^{\alpha} a$.
\end{definition}

\begin{definition} A semigroup with left cancellation is a semigroup
$\alpha$ such that for all $a,b,c : \alpha$
$a *^{\alpha} b = a *^{\alpha} c => b = c$.
\end{definition}

\begin{definition} A semigroup with right cancellation is a semigroup
$\alpha$ such that for all $a,b,c : \alpha$
$a *^{\alpha} b = c *^{\alpha} b => a = c$.
\end{definition}

\begin{signature} A type with one is a type.
\end{signature}
\begin{signature} Assume $\alpha$ is a type with one. $1^{\alpha}$ is an
element of $\alpha$.
\end{signature}

\begin{definition} A monoid is a semigroup $\alpha$ such that $\alpha$ is a type
with one and
$\forall a : \alpha 1^{\alpha} *^{\alpha} a = a$ and
$\forall a : \alpha a *^{\alpha} 1^{\alpha} = a$.
\end{definition}

\begin{definition} A commutative monoid is a monoid that
is a commutative semigroup.
\end{definition}

\begin{signature} A type with inverses is a type.
\end{signature}
\begin{signature} Assume $\alpha$ is a type with inverses and $a : \alpha$.
$a^{-1,\alpha}$ is an element of $\alpha$.
\end{signature}

\begin{definition} A group is a monoid $\alpha$ such that $\alpha$ is a type 
with inverses and 
for all $a$ : $\alpha a^{-1,\alpha} *^{\alpha} a = 1^{\alpha}$.
\end{definition}

\begin{definition} A commutative group is a group that is a commutative
monoid.
\end{definition}

\begin{lemma}[mul left comm] Let $\alpha$ be a commutative semigroup.
Then for all $a,b,c : \alpha$ 
$a *^{\alpha} (b *^{\alpha} c) = b *^{\alpha} (a *^{\alpha} c)$.
\end{lemma}

\begin{lemma}[mul right comm] Let $\alpha$ be a commutative semigroup.
Then for all $a,b,c : \alpha$ 
$a *^{\alpha} (b *^{\alpha} c) = a *^{\alpha} (c *^{\alpha} b)$.
\end{lemma}

\begin{lemma}[mul left cancel iff] Let $\alpha$ be a 
semigroup with left cancellation. Then for all $a,b,c : \alpha$ 
$a *^{\alpha} b = a *^{\alpha} c <=> b = c$.
\end{lemma}

\begin{lemma}[mul right cancel iff] Let $\alpha$ be a 
semigroup with right cancellation. Then for all $a,b,c : \alpha$ 
$b *^{\alpha} a = c *^{\alpha} a <=> b = c$.
\end{lemma}

Let $\alpha$ denote a group.

\begin{lemma}[inv mul cancel left] For all $a,b : \alpha$ 
$a^{-1,\alpha} *^{\alpha} (a *^{\alpha} b) = b$.
\end{lemma}

\begin{lemma}[inv mul cancel right] For all $a,b : \alpha$ 
$a *^{\alpha} (b^{-1,\alpha} *^{\alpha} b) = a$.
\end{lemma}

\begin{lemma}[inv eq of mul eq one] Let $a, b : \alpha$ and
$a *^{\alpha} b = 1^{\alpha}$. Then $a^{-1,\alpha} = b$.
\end{lemma}

\begin{lemma}[one inv] $(1^{\alpha})^{-1,\alpha} = 1^{\alpha}$.
\end{lemma}

\begin{lemma}[inv inv] Let $a : \alpha$. Then 
$(a^{-1,\alpha})^{-1,\alpha} = a$.
\end{lemma}

\begin{lemma}[mul right inv] Let $a : \alpha$. Then 
$a *^{\alpha} a^{-1,\alpha} = 1^{\alpha}$.
\end{lemma}

\begin{lemma}[inv inj] Let $a,b : \alpha$ and $a^{-1,\alpha} = b^{-1,\alpha}$.
Then $a = b$.
\end{lemma}

\begin{lemma}[group mul left cancel] Let $a,b,c : \alpha$ and 
$a *^{\alpha} b = a *^{\alpha} c$. Then $b = c$.
\end{lemma}

\begin{lemma}[group mul right cancel] Let $a,b,c : \alpha$ and 
$a *^{\alpha} b = c *^{\alpha} b$. Then $a = c$.
\end{lemma}

\begin{proof} $a = (a *^{\alpha} b) *^{\alpha} b^{-1,\alpha}
= (c *^{\alpha} b) *^{\alpha} b^{-1,\alpha} = c$.
\end{proof}

\begin{lemma}[mul inv cancel left] Let $a,b : \alpha$. Then 
$a *^{\alpha} (a^{-1,\alpha} *^{\alpha} b) = b$.
\end{lemma}

\begin{lemma}[mul inv cancel right] Let $a,b : \alpha$. Then 
$(a *^{\alpha} b) *^{\alpha} b^{-1,\alpha} = a$.
\end{lemma}


\begin{lemma}[mul inv rev] Let $a,b : \alpha$. Then
$(a *^{\alpha} b)^{-1,\alpha} = b^{-1,\alpha} *^{\alpha} a^{-1,\alpha}$.
\end{lemma}
\begin{proof} 
$(a *^{\alpha} b) *^{\alpha} (b^{-1,\alpha} *^{\alpha} a^{-1,\alpha})
= 1^{\alpha}$.
\end{proof}

\begin{lemma}[eq inv of eq inv] Let $a,b : \alpha$ and $a = b^{-1,\alpha}$.
Then $b = a^{-1,\alpha}$.
\end{lemma}

\begin{lemma}[eq inv of mul eq one] Let $a,b : \alpha$ and 
$a *^{\alpha} b = 1^{\alpha}$. Then $a = b^{-1,\alpha}$.
\end{lemma}

\begin{lemma}[eq mul inv of mul eq] Let $a,b,c : \alpha$ and 
$a *^{\alpha} c = b$. Then $a = b *^{\alpha} c^{-1,\alpha}$.
\end{lemma}

\begin{lemma}[eq inv mul of mul eq] Let $a,b,c : \alpha$ and 
$b *^{\alpha} a = c$. Then $a = b^{-1,\alpha} *^{\alpha} c$.
\end{lemma}

\begin{lemma}[inv mul eq of eq mul] Let $a,b,c : \alpha$ and
$b = a *^{\alpha} c$. Then $a^{-1,\alpha} *^{\alpha} b = c$.
\end{lemma}

\begin{lemma}[mul inv eq of eq mul] Let $a,b,c : \alpha$ and 
$a = c *^{\alpha} b$. Then $a *^{\alpha} b^{-1,\alpha} = c$.
\end{lemma}

\begin{lemma}[eq mul of mul inv eq] Let $a,b,c : \alpha$ and 
$a *^{\alpha} c^{-1,\alpha} = b$. Then $a = b *^{\alpha} c$.
\end{lemma}

\begin{lemma}[eq mul of inv mul eq] Let $a,b,c : \alpha$ and 
$b^{-1,\alpha} *^{\alpha} a = c$. Then $a = b *^{\alpha} c$.
\end{lemma}

\begin{lemma}[mul eq of eq inv mul] Let $a,b,c : \alpha$ and 
$b = a^{-1,\alpha} *^{\alpha} c$. Then $a *^{\alpha} b = c$.
\end{lemma}

\begin{lemma}[mul eq of eq mul inv] let $a,b,c : \alpha$ and
$a = c *^{\alpha} b^{-1,\alpha}$. Then $a *^{\alpha} b = c$.
\end{lemma}

\begin{lemma}[mul inv] Let $\alpha$ be a commutative group. 
Let $a,b : \alpha$. Then 
$(a *^{\alpha} b)^{-1,\alpha} = a^{-1,\alpha} *^{\alpha} b^{-1,\alpha}$.
\end{lemma}

\end{forthel}
\end{document}

