\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb, mathtools}

\usepackage{../lib/tex/naproche}

\title{König's Theorem}
\author{Steffen Frerix and Peter Koepke\\
University of Bonn}
\date{2018 - 2021}

\newcommand{\card}[1]{\operatorname{card}(#1)}

\begin{document}

  \maketitle

König's Theorem is an important set-theoretical result about the
arithmetic of cardinals. It was proved by Julius König in 1905
(König, Julius: Zum Kontinuumsproblem, Mathematische Annalen 60 (1905), 177–180).
The proof is reminiscent of Cantor's diagonal argument for
proving that $\kappa < 2^\kappa$.

The formalization begins with a minimal axiomatic setup on sets, functions
and cardinals, sufficient for the proof of the theorem.

On an older laptop
with an Intel Pentium N3710 processor \Naproche takes
around 35 seconds to check this text.

\section{ForTheL Setup}

\begin{forthel}
    [synonym cardinal/-s]
    [synonym sequence/-s]
    [read ZFC.ftl]
    [read macros.ftl]
\end{forthel}


\section{Sets and Functions}

\begin{forthel}
    Let $M,N$ denote sets.

    \begin{definition}
     A subset of $M$ is a set $N$ such that every element
     of $N$ is an element of $M$.
    \end{definition}

    \begin{definition}
      $M \setminus N = \{x \in M | x$ is not an element of $N\}$.
    \end{definition}

    Let $f$ denote a function.

    \begin{definition}
      Assume $M$ is a subset of the domain of $f$.
      $f[M] = \{f(x) | x$ is an element of $M\}$.
    \end{definition}

    Let the image of $f$ stand for $f[\dom(f)]$.

    \begin{axiom}
      Assume $M$ is a subset of the domain of $f$.
      Then $f[M]$ is a set.
    \end{axiom}

    \begin{axiom}
      $f$ is setsized.
    \end{axiom}
\end{forthel}

  \section{Cardinals}

\begin{forthel}
    \begin{signature}
      A cardinal is a set.
    \end{signature}

    Let $A,B,C$ denote cardinals.
    \begin{signature}
      $A < B$ is an atom.
    \end{signature}

    Let $A$ is less than $B$ stand for $A < B$.
    Let $A \leq B$ stand for $A = B$ or $A < B$.

    \begin{axiom}
      $A < B < C \implies A < C$.
    \end{axiom}

    \begin{axiom}
      $A < B$ or $B < A$ or $B = A$.
    \end{axiom}

    \begin{signature}
      The cardinality of $M$ is a cardinal.
    \end{signature}

    Let $\card{M}$ denote the cardinality of $M$.

    \begin{axiom}
      $\card{A} = A$.
    \end{axiom}

    \begin{axiom}[ImageCard]
      Assume $M$ is a subset of $Dom(f)$. $\card{f[M]} \leq \card{M}$.
    \end{axiom}

    \begin{axiom}
      Assume the cardinality of $N$ is less than the cardinality of $M$.
Then $M \setminus N$ has an element.
    \end{axiom}

    \begin{axiom}[SurjExi]
      Assume $\card{M} \leq \card{N}$. Assume $M$ has an element. There exists a
      function $f$ such that $N$ is the domain of $f$ and $M$ is the image of $f$.
    \end{axiom}

\end{forthel}

\section{Sums and Products of cardinals}

\newcommand{\SumSet}[2]{{\dot{\bigcup}}_{i \in #2} #1_{i}}
\newcommand{\Sum}[2]{{\sum}_{i \in #2} #1_{i}}
\newcommand{\ProdSet}[2]{{\bigtimes}_{i \in #2} #1_{i}}
\newcommand{\Prod}[2]{{\prod}_{i \in #2} #1_{i}}

  \begin{forthel}
    Let $D$ denote a set.

    \begin{definition}
      A sequence of cardinals on $D$ is a function $\kappa$
      such that $\dom(\kappa) = D$
      and $\kappa(i)$ is a cardinal for every element $i$ of $D$.
    \end{definition}

Let $f_{i}$ stand for $f(i)$.

    \begin{signature}
      Let $\kappa$ be a sequence of cardinals on $D$.
     $\SumSet{\kappa}{D}$ is a set.
    \end{signature}

    \begin{axiom}[SumDef]
      Let $\kappa$ be a sequence of cardinals on $D$. $\SumSet{\kappa}{D} =
      \{(n,i) | i$ is an element of $D$ and $n$ is an element
      of $\kappa_{i}\}$.
    \end{axiom}

    \begin{axiom}
      Let $\kappa$ be a sequence of cardinals on $D$.
      Then $\SumSet{\kappa}{D}$ is a set.
    \end{axiom}

    \begin{definition}
      Let $\kappa$ be a sequence of cardinals on $D$.
      $\Sum{\kappa}{D} = \card{\SumSet{\kappa}{D}}$.
    \end{definition}

    \begin{signature}
      Let $\kappa$ be a sequence of cardinals on $D$.
      $\ProdSet{\kappa}{D}$ is a set.
    \end{signature}

    \begin{axiom}[ProdDef]
      Let $\kappa$ be a sequence of cardinals on $D$. $\ProdSet{\kappa}{D} =
      \{$ function $f | \dom(f) = D \wedge (f(i)$ is an element of
      $\kappa_{i}$ for every
      element $i$ of $D)\}$.
    \end{axiom}

    \begin{axiom}
      Let $\kappa$ be a sequence of cardinals on $D$.
      Then $\ProdSet{\kappa}{D}$ is a set.
    \end{axiom}

    \begin{definition}
      Let $\kappa$ be a sequence of cardinals on $D$.
      $\Prod{\kappa}{D} = \card{\ProdSet{\kappa}{D}}$.
    \end{definition}
\end{forthel}

König's Theorem requires some form of the axiom of choice.
Currently choice is built into Naproche by the {\em choose}
construct in function definitions.

The axiom of choice is also required to show that products
of non-empty factors are themselves non-empty:

\begin{forthel}

    \begin{lemma}[Choice]
      Let $\lambda$ be a sequence of cardinals on $D$.
      Assume that $\lambda_{i}$
     has an element for every element $i$ of $D$.
     Then $\ProdSet{\lambda}{D}$ has an element.
    \end{lemma}
    \begin{proof}
      Define $f(i) =$ choose an element $v$ of
      $\lambda_{i}$ in $v$ for $i$ in $D$.
     Then $f$ is an element of $\ProdSet{\lambda}{D}$.
    \end{proof}

\end{forthel}


\section{König's theorem}

\begin{forthel}
    \begin{theorem}
      Let $\kappa, \lambda$ be sequences of cardinals on $D$.
     Assume that for every element $i$ of $D$
     $\kappa_{i} < \lambda_{i}$.
     Then $$\Sum{\kappa}{D} < \Prod{\lambda}{D}.$$
    \end{theorem}
    \begin{proof}
      Proof by contradiction. Assume the contrary.
      Then $$\Prod{\lambda}{D} \leq \Sum{\kappa}{D}.$$
      Take a function $G$ such that $\SumSet{\kappa}{D}$ is the domain of $G$
      and $\ProdSet{\lambda}{D}$ is the image of $G$.
      Indeed $\ProdSet{\lambda}{D}$ has an element.

      Define

      $\Delta(i) =
     \{G((n,i))(i) | n$ is an element of $\kappa_{i}\}$

      for $i$ in $D$.

      For every element $f$ of $\ProdSet{\lambda}{D}$
      for every element $i$ of $D$
      $f(i)$ is an element of $\lambda_{i}$.
      For every element $i$ of $D$ $\lambda_{i}$ is a set.
      For every element $i$ of $D$ for every element $d$ of $\Delta(i)$
      we have $d\in \lambda_{i}$.
      For every element $i$ of $D$ $\Delta(i)$ is a set.

     (1) For every element $i$ of $D$
      $\card{\Delta(i)} < \lambda_{i}$.

      Proof.
        Let $i$ be an element of $D$.
        Define $F(n) = G((n,i))(i)$ for $n$ in $\kappa_{i}$.
        Then $F[\kappa(i)] = \Delta(i)$.
      end.

      Define
      $f(i) =$ choose an element $v$ of
      $\lambda_{i} \setminus \Delta(i)$ in
      $v$ for $i$ in $D$. Then $f$ is an element of
      $\ProdSet{\lambda}{D}$. Take an element $j$ of $D$ and an
      element $m$ of $\kappa_{j}$ such that $G((m,j)) = f$.
      $G((m,j))(j)$ is an element of $\Delta(j)$ and $f(j)$
      is not an element of $\Delta(j)$. Contradiction.
    \end{proof}
  \end{forthel}




\end{document}
