\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{../lib/tex/forthel}

\title{Maximum modulus principle}
\author{Steffen Frerix}
\date{2018}

\begin{document}
  \pagenumbering{gobble}

  \maketitle

  \begin{forthel}
    [synonym number/-s]

    Let the domain of $f$ stand for $Dom(f)$.
    Let $z$ is in $M$ stand for $z$ is an element of $M$.
    Let $M$ contains $z$ stand for $z$ is in $M$.
    Let $z << M$ stand for $z$ is in $M$.

    Let $f$ denote a function.
    Let $M$ denote a set.

    \begin{definition}
      A subset of $M$ is a set $N$ such that every element of $N$ is an element of $M$.
    \end{definition}

    \begin{definition}
      Assume $M$ is a subset of the domain of $f$. $f^(M) = { f(x) | x << M }$.
    \end{definition}

    \begin{axiom}
      Assume $M$ is a subset of the domain of $f$. $f^(M)$ is a set.
    \end{axiom}

    \begin{signature}
      A complex number is a notion.
    \end{signature}

    Let $z,w$ denote complex numbers.

    \begin{axiom}
      Every element of $Dom(f)$ is a complex number and for every element $z$ of $Dom(f)$ $f(z)$ is a complex number.
    \end{axiom}

    \begin{axiom}
      Every element of $M$ is a complex number.
    \end{axiom}

    \begin{signature}
      A real number is a notion.
    \end{signature}

    Let $x,y$ denote real numbers.

    \begin{axiom}
      $x$ is setsized.
    \end{axiom}

    \begin{signature}
      $|z|$ is a real number.
    \end{signature}

    \begin{signature}
      $x$ is positive is an atom.
    \end{signature}

    Let $eps, delta$ denote positive real numbers.

    \begin{signature}
      $x < y$ is an atom.
    \end{signature}

    Let $x =< y$ stand for $x = y$ or $x < y$.

    \begin{axiom}
      $x < y =>$ not $y < x$.
    \end{axiom}

    \begin{signature}
      $f$ is holomorphic is an atom.
    \end{signature}

    \begin{signature}
      $Ball(eps,z)$ is a set that contains $z$.
    \end{signature}

    \begin{axiom}
      Assume $x$ is element of $Ball(eps, z)$. Then $x$ is a real number.
    \end{axiom}

    \begin{lemma}
      Assume $x$ is an element of $Ball(eps, z)$. Then $x$ is setsized.
    \end{lemma}

    \begin{axiom}
      $|z| < |w|$ for some element $w$ of $Ball(eps,z)$.
    \end{axiom}

    \begin{definition}
      $M$ is open iff for every element $z$ of $M$ there exists $eps$ such that $Ball(eps,z)$ is a subset of $M$.
    \end{definition}

    \begin{axiom}
      $Ball(eps,z)$ is open.
    \end{axiom}

    \begin{definition}
      A local maximal point of $f$ is an element $z$ of the domain of $f$ such that there exists $eps$ such that $Ball(eps,z)$ is a subset of the domain of $f$ and $|f(w)| =< |f(z)|$ for every element $w$ of $Ball(eps,z)$.
    \end{definition}

    \begin{definition}
      Let $U$ be a subset of the domain of $f$. $f$ is constant on $U$ iff there exists $z$ such that $f(w) = z$ for every element $w$ of $U$.
    \end{definition}

    Let $f$ is constant stand for $f$ is constant on the domain of $f$.

    \begin{axiom}
      Assume $f$ is holomorphic and $Ball(eps,z)$ is a subset of the domain of $f$. If $f$ is not constant on $Ball(eps,z)$ then $f^(Ball(eps,z))$ is open.
    \end{axiom}

    \begin{signature}
      A region is an open set.
    \end{signature}

    \begin{axiom}[IdentityTheorem]
      Assume $f$ is holomorphic and the domain of $f$ is a region. Assume that $Ball(eps,z)$ is a subset of the domain of $f$. If $f$ is constant on $Ball(eps,z)$ then $f$ is constant.
    \end{axiom}

    \begin{proposition}[Maximumprinciple]
      Assume $f$ is holomorphic and the domain of $f$ is a region. If $f$ has a local maximal point then $f$ is constant.
    \end{proposition}
    \begin{proof}
      Let $z$ be a local maximal point of $f$. Take $eps$ such that $Ball(eps,z)$ is a subset of $Dom(f)$ and $|f(w)| =< |f(z)|$ for every element $w$ of $Ball(eps,z)$.

      Let us show that $f$ is constant on $Ball(eps,z)$.
      proof by contradiction.
        Assume the contrary. Then $f^(Ball(eps,z))$ is open. We can take delta such that $Ball(delta, f(z))$ is a subset of $f^(Ball(eps,z))$. Therefore there exists an element $w$ of $Ball(eps,z)$ such that $|f(z)| < |f(w)|$. Contradiction.
    	end.

      Hence $f$ is constant.
    \end{proof}
  \end{forthel}

\end{document}
