\documentclass[a4paper,draft]{amsproc}
\title{\textbf{Powerset}}

\date{}
\begin{document}

\theoremstyle{plain}
 \newtheorem{theorem}{Theorem}[section]
 \newtheorem{proposition}{Proposition}[section]
 \newtheorem{lemma}{Lemma}[section]
 \newtheorem{corollary}{Corollary}[section]
 \newtheorem{axiom}{Axiom}[section]
\theoremstyle{definition}
 \newtheorem{example}{Example}[section]
 \newtheorem{definition}{Definition}[section]
\theoremstyle{remark}
 \newtheorem{remark}{Remark}[section]
 \newtheorem{notation}{Notation}[section]
 \numberwithin{equation}{section}

% This is a snippet that lets us use quotation marks in order to write regular text in math mode.
% Naproche-SAD will ignore all quotation marks.
\let\ftlquotationmark"
\catcode`\"=\active
\def"#1"{\ifmmode\mathrel{\textrm{#1}}\else \ftlquotationmark #1\ftlquotationmark \fi}

\newenvironment{forthel}{}{}

\maketitle 
\section{First and only section}

\begin{theorem}
This is not parsed by forthel.
\end{theorem}

\begin{forthel}
[synonym subset/-s]

[synonym surject/-s]

% We add space in between lines in order to let this file produce a reasonable-looking
% latex pdf.
Let $M$ denote a set.

Let $f$ denote a function.

Let the value of $f$ at $x$ stand for $f[x]$.

Let $f$ is defined on $M$ stand for $Dom(f) = M$.

Let the domain of $f$ stand for $Dom(f)$.


\begin{axiom}
The value of $f$ at any element of the domain of $f$ is a set.
\end{axiom}

\begin{definition}[subset]
A subset of $M$ is a set N such that every element of N is an element of $M$.
\end{definition}

\begin{definition}
The powerset of $M$ is the set of subsets of $M$.
\end{definition}

\begin{definition}
$f$ surjects onto $M$ iff every element of $M$ is equal to the value of $f$ at some element of the domain of $f$.
\end{definition}

\begin{proposition}
No function that is defined on $M$ surjects onto the powerset of $M$.
\end{proposition}
\begin{proof}
Proof by contradiction.
Assume the contrary. Take a function $f$ that is defined on $M$ and surjects onto the powerset of $M$.

Define $N = \{ x \in M | x "is not an element of" f[x] \}$.
Then $N$ is not equal to the value of $f$ at any element of $M$.
Contradiction.
\end{proof}

\begin{theorem}
The value of $f$ at any element of the domain of $f$ is a set.
\end{theorem}

\end{forthel}

\end{document}