\documentclass{stex}
\libinput{preamble}
\begin{document}

\section{Sets and Classes}

Some basic properties of sets and classes are built into \Naproche{}
or are explicitly postulated in the \path{preliminaries.ftl.tex} file.
We shall use these to express the infinitude of the collection of
prime natural numbers.

Arbitrary collections of objects can be defined naturally in ForTheL:

\begin{forthel}
\begin{definition} $\mathbb{N}$ is the
collection of all natural numbers.
\end{definition}
\end{forthel}

This translates as
\begin{verbatim}
forall v0 ((HeadTerm :: v0 = \mathbb{N}) iff
(aClass(v0) and forall v1
(aElementOf(v1,v0) iff (aNaturalNumber(v1) and aObject(v1)))))
\end{verbatim}

The \verb+aObject(v1)+ is part of the collection construct and ensures
that only objects will be elements of the collection. Such collections
are automatically registered as {\em classes} (\verb+aClass(v0)+).
There are a few grammatical extensions and variants of this construct:
in particular one could form the collection of all
[terms] such that [condition], where [condition] is an arbitrary ForTheL
statement.

Classes can also be defined by abstraction terms, i.e., we could
have defined alternatively:

\begin{definition}
$\mathbb{N} = \{n | n$ is a natural number$\}$.
\end{definition}

To work freely with the class of natural numbers one wants it be a
{\em mathematical object} itself, i.e., a {\em set}. The axioms so far do not
imply this, as one can check by trying to prove
``$\mathbb{N}$ is a set.''
So we postulate this as an axiom which is (equivalent to) the
well-known axiom of infinity in set theory:

\begin{forthel}
\begin{axiom} $\mathbb{N}$ is a set.
\end{axiom}
\end{forthel}

We define the set of prime numbers so that we can prove its infinity.

\begin{forthel}
\begin{definition}
$\mathbb{P} = \{n | n$ is a prime natural number$\}$.
\end{definition}
\end{forthel}
The axiom of separation from \path{preliminaries.ftl.tex} yields:
\begin{forthel}
\begin{lemma}
$\mathbb{P}$ is a set.
\end{lemma}
\end{forthel}
\noindent since $\mathbb{P}$ is a subclass of the {\em set} $\mathbb{N}$.

\begin{exercise}
Form the collection of all even numbers or of all square numbers and
prove that these are sets.
\end{exercise}

\begin{exercise}
Prove that the class $\{ x | x$ is a set and $x \notin x \}$ is \textit{not}
a set.
\end{exercise}

\end{document}
