\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{../../../meta-inf/lib/naproche}
\input{../../lib/libraries}
\input{../../lib/foundations}
\input{../../lib/arithmetics}
\begin{document}
  \begin{imports}
    \begin{forthel}
      %[prove off][check off]
      [readtex \path{libraries/source/foundations/finite-and-infinite-classes.ftl.tex}]
      %[prove on][check on]
    \end{forthel}
  \end{imports}


  \section*{Countable and Uncountable Classes}


  \subsection*{Countably Infinite Classes}

  \begin{forthel}
    \begin{definition}\printlabel{FOUNDATIONS_14_6249029537103872}
      Let $X$ be a class.
      $X$ is countably infinite iff $X$ is equinumerous to $\Nat$.
    \end{definition}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}\printlabel{FOUNDATIONS_14_803449379749888}
      Let $X, Y$ be classes.
      If $X$ is countably infinite and $Y$ is equinumerous to $X$ then $Y$ is countably infinite.
    \end{proposition}
    \begin{proof}
      Assume that $X$ is countably infinite and $Y$ is equinumerous to $X$.
      Take a bijection $f$ between $\Nat$ and $X$ and a bijection $g$ between $X$ and $Y$.
      Then $g \circ f$ is a bijection between $\Nat$ and $Y$ (by \printref{FOUNDATIONS_08_6435206693126144}).
      Indeed $X, Y$ are classes.
      Hence $Y$ is countably infinite.
    \end{proof}
  \end{forthel}


  \subsection*{Countable Classes}

  \begin{forthel}
    \begin{definition}\printlabel{FOUNDATIONS_14_5412969443753984}
      Let $X$ be a class.
      $X$ is countable iff $X$ is finite or $X$ is countably infinite.
    \end{definition}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}\printlabel{FOUNDATIONS_14_4182588499427328}
      Let $X, Y$ be classes.
      If $X$ is countable and $Y$ is equinumerous to $X$ then $Y$ is countable.
    \end{proposition}
    \begin{proof}
      Assume that $X$ is countable and $Y$ is equinumerous to $X$.
      If $X$ is finite then $Y$ is finite.
      If $X$ is countably infinite then $Y$ is countably infinite.
      Hence $Y$ is countable.
    \end{proof}
  \end{forthel}


  \subsection*{Uncountable Classes}

  \begin{forthel}
    \begin{definition}\printlabel{FOUNDATIONS_14_2411928395710464}
      Let $X$ be a class.
      $X$ is uncountable iff $X$ is not countable.
    \end{definition}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}\printlabel{FOUNDATIONS_14_1073385358491648}
      Let $X, Y$ be classes.
      If $X$ is uncountable and $Y$ is equinumerous to $X$ then $Y$ is uncountable.
    \end{proposition}
    \begin{proof}
      Assume that $Y$ is equinumerous to $X$.
      If $Y$ is countable then $X$ is countable.
      Hence if $X$ is uncountable then $Y$ is uncountable.
    \end{proof}
  \end{forthel}
\end{document}
