\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{../../../meta-inf/lib/naproche}
\input{../../lib/libraries}
\input{../../lib/foundations}
\input{../../lib/arithmetics}
\begin{document}
  \begin{imports}
    \begin{forthel}
      %[prove off][check off]
      [readtex \path{libraries/source/foundations/segments-of-natural-numbers.ftl.tex}]
      [readtex \path{libraries/source/foundations/equinumerosity.ftl.tex}]
      %[prove on][check on]
    \end{forthel}
  \end{imports}


  \section*{Finite and Infinite Classes}

  \subsection*{Finite Classes}

  \begin{forthel}
    \begin{definition}[id=FOUNDATIONS_14_3512046897512410,printid]
      Let $X$ be a class and $n$ be a natural number.
      $X$ has $n$ elements iff $X$ is equinumerous to $\{ 1, \dots, n \}$.
    \end{definition}
  \end{forthel}

  \begin{forthel}
    \begin{definition}[id=FOUNDATIONS_14_3694156977274880,printid]
      Let $X$ be a class.
      $X$ is finite iff there exists a natural number $n$ such that $X$ has $n$ elements.
    \end{definition}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_14_3929085203972096,printid]
      Let $X, Y$ be classes.
      If $X$ is finite and $Y$ is equinumerous to $X$ then $Y$ is finite.
    \end{proposition}
    \begin{proof}
      Assume that $X$ is finite and $Y$ is equinumerous to $X$.
      Take a natural number $n$ and a bijection $f$ between $\{ 1, \dots, n \}$ and $X$ and a bijection $g$ between $X$ and $Y$.
      Then $g \circ f$ is a bijection between $\{ 1, \dots, n \}$ and $Y$ (by \printref{FOUNDATIONS_08_6435206693126144}).
      Indeed $X, Y$ are classes.
      Hence $Y$ is finite.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_14_5132547854597502,printid]
      Let $X$ be a class.
      $X$ has $0$ elements iff $X = \emptyset$.
    \end{proposition}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_14_6812054297034125,printid]
      Let $X$ be a class.
      $X$ has $1$ element iff $X = \set{a}$ for some object $a$.
    \end{proposition}
    \begin{proof}
      Case $X$ has $1$ element.
        Take a bijection $f$ between $\{ 1, \dots, 1 \}$ and $X$.
        We have $\{ 1, \dots, 1 \} = \set{1}$.
        Hence $X = \set{f(1)}$.
      End.

      Case $X = \set{a}$ for some object $a$.
        Consider an object $a$ such that $X = \set{a}$.
        Define $f(x) = 1$ for $x \in \set{a}$.
      Then $f$ is a bijection between $\set{a}$ and $\{ 1, \dots, 1 \}$.
      End.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_14_3468912675458910,printid]
      Let $X$ be a class.
      $X$ has $2$ elements iff $X = \set{a, b}$ for some distinct objects $a, b$.
    \end{proposition}
    \begin{proof}
      Case $X$ has $2$ elements.
        Take a bijection $f$ between $\{ 1, \dots, 2 \}$ and $X$.
        We have $\{ 1, \dots, 2 \} = \set{1, 2}$.
        Hence $X = \set{f(1), f(2)}$.
        We have $f(1) \neq f(2)$.
      End.

      Case $X = \set{a, b}$ for some distinct objects $a, b$.
        Consider distinct objects $a, b$ such that $X = \set{a, b}$.
        Define \[f(x) =
          \begin{cases}
            1 & : x = a \\
            2 & x = b
          \end{cases}\]
        for $x \in \set{a, b}$.
        Then $f$ is a bijection between $\set{a, b}$ and $\{ 1, \dots, 2 \}$.
        Indeed $f$ is injective and surjective onto $\{ 1, \dots, 2 \}$.
      End.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_14_0615204230800975,printid]
      Let $n$ be a natural number and $X$ be a class that has $n$ elements and $a$ be an object such that $a \notin X$.
      Then $X \cup \set{a}$ has $n + 1$ elements.
    \end{proposition}
    \begin{proof}
      Take a bijection $f$ between $X$ and $\{ 1, \dots, n \}$.
      Define \[g(x) =
        \begin{cases}
          f(x)  & : x \in X \\
          n + 1 & : x = a
        \end{cases}\]
      for $x \in X \cup \set{a}$.

      (1) $g$ is a map from $X \cup \set{a}$ to $\{ 1, \dots, n + 1 \}$. \\
      Indeed we can show that $g(x) \in \{ 1, \dots, n + 1 \}$ for all $x \in X \cup \set{a}$.
        Let $x \in X \cup \set{a}$.
        If $x \in X$ then $g(x) \in \{ 1, \dots, n \}$.
        If $x = a$ then $g(x) = n + 1$.
      End.

      (2) $g$ is injective. \\
      Proof.
        Let $x, y \in \dom(g)$.
        Assume $x \neq y$.
        
        Case $x, y \in X$. Obvious.

        Case $x \in X$ and $y = a$. Obvious.

        Case $x = a$ and $y \in X$. Obvious.
      Qed.

      (3) $g$ is surjective onto $\{ 1, \dots, n + 1 \}$.
      Indeed we can show that for every $k \in \{1, \dots, n + 1 \}$ there exists an $x \in \dom(g)$ such that $k = g(x)$. \\
      Proof.
        Let $k \in \{ 1, \dots, n + 1 \}$.

        Case $k \leq n$.
          Then $k \in \{ 1, \dots, n \}$.
          Hence we can take a $x \in X$ such that $k = f(x)$.
        End.

        Case $k = n + 1$.
          Then $k = g(a)$.
        End.
      Qed.

      Hence $g$ is a bijection between $X \cup \set{a}$ and $\{ 1, \dots, n + 1 \}$.
    \end{proof}
  \end{forthel}


  \subsection*{Infinite Classes}

  \begin{forthel}
    \begin{definition}[id=FOUNDATIONS_14_6612510618681344,printid]
      Let $X$ be a class.
      $X$ is infinite iff $X$ is not finite.
    \end{definition}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_14_5814530911240192,printid]
      Let $X, Y$ be classes.
      If $X$ is infinite and $Y$ is equinumerous to $X$ then $Y$ is infinite.
    \end{proposition}
    \begin{proof}
      Assume that $Y$ is equinumerous to $X$.
      If $Y$ is finite then $X$ is finite.
      Hence if $X$ is infinite then $Y$ is infinite.
    \end{proof}
  \end{forthel}
\end{document}
