\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{../../../meta-inf/lib/naproche}
\input{../../lib/libraries}
\input{../../lib/foundations}
\begin{document}
  \begin{imports}
    \begin{forthel}
      %[prove off][check off]
      [readtex \path{libraries/source/foundations/classes.ftl.tex}]
      %[prove on][check on]
    \end{forthel}
  \end{imports}


  \section*{Ordered Pairs and Cartesian Products}

  \subsection*{Pairs}

  \begin{forthel}
    \begin{axiom}[id=FOUNDATIONS_04_8464577431863296,printid]
      Let $a, a', b, b'$ be objects.
      If $(a, b) = (a', b')$ then $a = a'$ and $b = b'$.
    \end{axiom}
  \end{forthel}

  \begin{forthel}
    \begin{definition}[id=FOUNDATIONS_04_4782386822774784,printid]
      A pair is an object $p$ such that $p = (a, b)$ for some objects $a, b$.
    \end{definition}

    Let an ordered pair stand for a pair.
  \end{forthel}


  \subsection*{Cartesian Products}

  \begin{forthel}
    \begin{definition}[id=FOUNDATIONS_04_2877806274936832,printid]
      Let $A, B$ be classes.
      The Cartesian product of $A$ and $B$ is $\{ (a, b) \mid a \in A$ and $b \in B \}$.
    \end{definition}

    Let the direct product of $A$ and $B$ stand for  the Cartesian product of $A$ and $B$.
    Let $A \times B$ stand for the Cartesian product of $A$ and $B$.
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_04_1581118511906816,printid]
      Let $A, B$ be classes and $a, b$ be objects.
      Then $(a, b) \in A \times B$ iff $a \in A$ and $b \in B$.
    \end{proposition}
    \begin{proof}
      Case $(a, b) \in A \times B$.
        We can take $a' \in A$ and $b' \in B$ such that $(a, b) = (a', b')$.
        Then $a = a'$ and $b = b'$.
        Hence $a \in A$ and $b \in B$.
      End.

      Case $a \in A$ and $b \in B$.
        $a$ and $a$ are objects.
        Hence $(a, b)$ is an object.
        Therefore $(a, b) \in A \times B$.
      End.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_04_2198552029691904,printid]
      Let $A, B$ be classes.
      Then $A \times B$ is empty iff $A$ is empty or $B$ is empty.
    \end{proposition}
    \begin{proof}
      Case $A \times B$ is empty.
        Assume that $A$ and $B$ are nonempty.
        Then we can take an element $a$ of $A$ and an element $b$ of $B$.
        Then $(a, b) \in A \times B$.
        Contradiction.
      End.

      Case $A$ is empty or $B$ is empty.
        Assume that $A \times B$ is nonempty.
        Then we can take an element $c$ of $A \times B$.
        Then $c = (a, b)$ for some $a \in A$ and some $b \in B$.
        Hence $A$ and $B$ are nonempty.
        Contradiction.
      End.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_04_7971087096741888,printid]
      Let $a, b$ be objects.
      Then $\set{a} \times \set{b} = \set{(a, b)}$.
    \end{proposition}
    \begin{proof}
      Let us show that $\set{a} \times \set{b} \subseteq \set{(a, b)}$.
        Let $c \in \set{a} \times \set{b}$.
        Take $a' \in \set{a}$ and $b' \in \set{b}$ such that $c = (a', b')$.
        We have $a' = a$ and $b' = b$.
        Hence $c = (a, b)$.
        Thus $c \in \set{(a, b)}$.
      End.

      Let us show that $\set{(a, b)} \subseteq \set{a} \times \set{b}$.
        Let $c \in \set{(a, b)}$.
        Then $c = (a, b)$.
        We have $a \in \set{a}$ and $b \in \set{b}$.
        Hence $c \in \set{a} \times \set{b}$.
      End.
    \end{proof}
  \end{forthel}

  \begin{forthel}
    \begin{proposition}[id=FOUNDATIONS_04_7456594440749056,printid]
      Let $A, A', B, B'$ be nonempty classes.
      If $A \times B = A' \times B'$ then $A = A'$ and $B = B'$.
    \end{proposition}
    \begin{proof}
      Assume $A \times B = A' \times B'$.

      (1) $A \subseteq A'$ and $B \subseteq B'$. \\
      Proof.
        Let $a \in A$ and $b \in B$.
        Then $(a,b) \in A \times B$.
        Hence $(a,b) \in A' \times B'$.
        Thus $a \in A'$ and $b \in B'$.
      Qed.

      (2) $A' \subseteq A$ and $B' \subseteq B$. \\
      Proof.
        Let $a \in A'$ and $b \in B'$.
        Then $(a,b) \in A' \times B'$.
        Hence $(a,b) \in A \times B$.
        Thus $a \in A$ and $b \in B$.
      Qed.
    \end{proof}
  \end{forthel}
\end{document}
