\documentclass{naproche-library}
\libinput{preamble}
\begin{document}
\IfInputref{}{\setsectionlevel{section}}
\begin{smodule}[title=The Cantor-Schröder-Bernstein Theorem]{cantor-schroeder-bernstein.ftl}
\importmodule[libraries/set-theory]{knaster-tarski.ftl}

\begin{theorem*}[forthel,title=Cantor-Schröder-Bernstein,id=cantor_schroeder_bernstein]
  Let $x,y$ be sets.
  $x$ and $y$ are equinumerous iff there exists a injective map from $x$ to $y$ and there exists an injective map from $y$ to $x$.
\end{theorem*}
\begin{proof}[forthel]
  Case $x$ and $y$ are equinumerous.
    Take a bijection $f$ between $x$ and $y$.
    Then $f^{-1}$ is a bijection between $y$ and $x$.
    Hence $f$ is an injective map from $x$ to $y$ and $f^{-1}$ is an
    injective map from $y$ to $x$.
  End.

  Case there exists an injective map from $x$ to $y$ and there exists an injective map from $y$ to $x$.
    Take an injective map $f$ from $x$ to $y$.
    Take an injective map $g$ from $y$ to $x$.
    We have $y \setminus f[a] \subseteq y$ for any $a \in \pow(x)$.

    (1) Define $h(a) = x \setminus g[y \setminus f[a]]$ for $a \in \pow(x)$.

    $h$ is a map from $\pow(x)$ to $\pow(x)$.
    Indeed $h(a)$ is a subset of $x$ for each subset $a$ of $x$.

    Let us show that $h$ preserves subsets.
      Let $u, v$ be subsets of $x$.
      Assume $u \subseteq v$.
      Then $f[u] \subseteq f[v]$.
      Hence $y \setminus f[v] \subseteq y \setminus f[u]$.
      Thus $g[y \setminus f[v]] \subseteq g[y \setminus f[u]]$.
      Indeed $y \setminus f[v]$ and $y \setminus f[u]$ are subsets of $y$.
      Therefore $x \setminus g[y \setminus f[u]] \subseteq x \setminus g[y \setminus f[v]]$.
      Consequently $h[u] \subseteq h[v]$.
    End.

    Hence we can take a fixed point $c$ of $h$ (by \nameref{knaster_tarski}).

    (2) Define $F(u) = f(u)$ for $u \in c$.

    We have $c = h(c)$ iff $x \setminus c = g[y \setminus f[c]]$.
    $g^{-1}$ is a bijection between $\range(g)$ and $y$.
    Thus $x \setminus c = g[y \setminus f[c]] \subseteq \range(g)$.
    Therefore $x \setminus c$ is a subset of $\dom(g^{-1})$.

    (3) Define $G(u) = g^{-1}(u)$ for $u \in x \setminus c$.

    $F$ is a bijection between $c$ and $\range(F)$.
    $G$ is a bijection between $x \setminus c$ and $\range(G)$.

    Define \[ H(u) =
      \begin{cases}
        F(u) & : u \in c \\
        G(u) & : u \notin c
      \end{cases} \]
    for $u \in x$.

    Let us show that $H$ is a map to $y$.
      $\dom(H)$ is a set and every value of $H$ is an object.
      Hence $H$ is a map.

      Let us show that every value of $H$ lies in $y$.
        Let $v$ be a value of $H$.
        Take $u \in x$ such that $H(u) = v$.
        If $u \in c$ then $v = H(u) = F(u) = f(u) \in y$.
        If $u \notin c$ then $v = H(u) = G(u) = g^{-1}(u) \in y$.
      End.
    End.

    (4) $H$ is surjective onto $y$.
    Indeed we can show that every element of $y$ is a value of $H$.
      Let $v \in y$.

      Case $v \in f[c]$.
        Take $u \in c$ such that $f(u) = v$.
        Then $F(u) = v$.
      End.

      Case $v \notin f[c]$.
        Then $v \in y \setminus f[c]$.
        Hence $g(v) \in g[y \setminus f[c]]$.
        Thus $g(v) \in x \setminus h(c)$.
        We have $g(v) \in x \setminus c$.
        Therefore we can take $u \in x \setminus c$ such that $G(u) = v$.
        Then $v = H(u)$.
      End.
    End.

    (5) $H$ is injective.
    Indeed we can show that for all $u, v \in x$ if $u \neq v$ then $H(u) \neq H(v)$.
      Let $u,v \in x$.
      Assume $u \neq v$.

      Case $u,v \in c$.
        Then $H(u) = F(u)$ and $H(v) = F(v)$.
        We have $F(u) \neq F(v)$.
        Hence $H(u) \neq H(v)$.
      End.

      Case $u,v \notin c$.
        Then $H(u) = G(u)$ and $H(v) = G(v)$.
        We have $G(u) \neq G(v)$.
        Hence $H(u) \neq H(v)$.
      End.

      Case $u \in c$ and $v \notin c$.
        Then $H(u) = F(u)$ and $H(v) = G(v)$.
        Hence $v \in g[y \setminus F[c]]$.
        We have $G(v) = g^{-1}(v) \in y \setminus F[c]$.
        Indeed $v = g(v')$ for some $v' \in y \setminus F[c]$.
        Thus $G(v) \neq F(u)$.
      End.

      Case $u \notin c$ and $v \in c$.
        Then $H(u) = G(u)$ and $H(v) = F(v)$.
        Hence $u \in g[y \setminus F[c]]$.
        We have $G(u) = g^{-1}(u) \in y \setminus F[c]$.
        Indeed $u = g(u')$ for some $u' \in y \setminus F[c]$.
        Thus $G(u) \neq F(v)$.
      End.
    End.

    Consequently $H$ is a bijection between $x$ and $y$ (by 4, 5).
    Therefore $x$ and $y$ are equinumerous.
  End.
\end{proof}
\end{smodule}
\end{document}
