\documentclass{stex}
\libinput{preamble}
\begin{document}
\title{The Drinker Paradox in \Naproche}
\author{Marcel Schütz}
\date{2023}
\maketitle
\begin{smodule}{drinker.ftl}
  \importmodule[naproche/examples/preliminaries]{everyday-ontology.ftl}

  \noindent The \emph{Drinker Paradox} is a principle of classical predicate logic popularised by the logician Raymond Smullyan in his 1978 book \textit{What Is the Name of this Book?} \cite{Smullyan1978} which can be stated as:

  \begin{quotation}
    \noindent There is someone in the pub such that, if he is drinking then everyone in the pub is drinking.
  \end{quotation}

  \begin{forthel}
    \begin{signature*}
      The pub is an object.
    \end{signature*}

    \begin{theorem*}[Drinker Paradox]\label{drinker_paradox}
      Assume that there is a person inside the pub.
      Then there is a person $P$ inside the pub such that if $P$ is drinking then every person inside the pub is drinking.
    \end{theorem*}
    \begin{proof}[ by case analysis]
      Case every person inside the pub is drinking.
        Choose a person $P$ inside the pub.
        Then $P$ is drinking and every person inside the pub is drinking.
        Hence if $P$ is drinking then every person inside the pub is drinking.
      End.

      Case there is a person inside the pub that is not drinking.
        Consider a person $P$ inside the pub that is not drinking.
        Then if $P$ is drinking then every person inside the pub is drinking.
      End.
    \end{proof}
  \end{forthel}
\end{smodule}
\printbibliography
\end{document}
